---
title: "Simulation-based inference - confidence intervals"
author: ""
date: ""
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

## Main ideas

- Understand sample statistic variability

- Bootstrap idea and how to use it

- Generating bootstrap confidence intervals

- Correctly interpret confidence intervals
  
# Packages

```{r packages}
library(tidyverse)
library(infer)
```

# Data

Consider 20 1-bedroom apartments that were randomly selected on 
Craigslist Manhattan from apartments listed as "by owner".

```{r data_manhattan}
manhattan <- read_csv("data/manhattan.csv")
```

Consider 3-10 day survival of mice randomized to various
neutron doses and streptomycin therapy or saline control.

```{r data_mice}
mice <- read_table("http://users.stat.ufl.edu/~winner/data/micerad.dat",
                   col_names = FALSE) %>% 
  rename(dose = X1, treat = X2, died = X3)
```

# Notes

## Variability of sample statistics

Suppose the following represents the population of IQ scores. There are only
1000 individuals in our population.

```{r population_iq_distribution}
population <- tibble(iq = rnorm(n = 1000, mean = 100, sd = 10))

population %>% 
  ggplot(aes(x = iq)) +
  geom_histogram(binwidth = 5, fill = "grey90", color = "blue") +
  labs(x = "IQ scores") +
  theme_bw()
```

Take a sample with `slice_sample()` and compute the mean.

```{r sample_iq_scores}
population %>% 
  slice_sample(n = 30) %>% 
  summarise(mean_iq = mean(iq))
```

Did everyone get the same value? If fact, each time you sample (run the above
code chunk) you will most likely get a slightly different mean.

The mean is not unique to this phenomenon. There will be variability for
all of the sample statistics we'll be using to produce confidence intervals.

If we want to construct a confidence interval for a population mean (or
any other parameter), we need to come up with a plausible range of values around 
our observed sample mean (statistic). This range will depend on how precise and 
how accurate our sample mean (statistic) is as an estimate of the population 
mean (parameter). Quantifying this requires a measurement of how much we would 
expect the sample mean (statistic) to vary from sample to sample.

## Confidence intervals via bootstrap simulation

To ensure reproducibility for your knitted document, set the random number
generation seed.

```{r set_seed}
set.seed(1902370)
```

### A confidence interval for the population mean - $\mu$

Create a 95% bootstrap confidence interval for the population *mean* price of
1-bedroom apartments in Manhattan.

First, identify the following:

- Population: 
- Parameter of interest:
- Sample:
- Sample size:

To create our confidence interval, we'll follow the four-step bootstrap
scheme outlined in the slides using functions from `infer`.

1. **Take a bootstrap sample** - a random sample taken with replacement 
   from the original sample, of the same size as the original sample.

2. **Calculate the bootstrap statistic** - a statistic such as mean, median,
   proportion, slope, etc. computed from the bootstrap samples.

3. **Repeat steps (1) and (2) many times to create a bootstrap distribution** - 
   a distribution of bootstrap statistics.

4. **Calculate the bounds of the XX% confidence interval** as the middle XX% 
   of the bootstrap distribution.
   
#### `infer`

First, `specify()` the variable of interest, in this case `rent`.

```{r specify_manhattan}
manhattan %>% 
  specify(response = rent)
```

Second, `generate()` a fixed number of bootstrap samples. Here we'll generate
10,000 bootstrap samples. Generally, the smaller your sample size, the more
bootstrap samples you'll want to generate. 

```{r generate_manhattan}
manhattan %>% 
  specify(response = rent) %>% 
  generate(reps = 10000, type = "bootstrap")
```

Finally, `calculate()` the relevant statistic for each bootstrap sample.

```{r calculate_manhattan}
boot_means <- manhattan %>% 
  specify(response = rent) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")

boot_means
```

Typically, you will do this all in one code chunk as is given in the chunk
named `calculate`. It is broken up here to better understand the steps.

Visualize `boot_means` with `ggplot()`.

```{r visualize_means_manhattan}
boot_means %>% 
  ggplot(aes(x = stat)) +
  geom_histogram(binwidth = 50)
```

Finally, compute the lower and upper bounds to form our interval estimate.
For a 95% confidence interval, these bounds occur at the 2.5% and 97.5% 
quantiles.

```{r ci_rent_dplyr}
boot_means %>% 
  summarise(
    lb = quantile(stat, 0.025),
    ub = quantile(stat, 0.975)
  )
```

Package `infer` also has a convenient function to do this: `get_ci()`. Use
whichever method you prefer.

```{r ci_rent_infer}
get_ci(boot_means, level = 0.95)
```

What does this mean?

**Interpretation**:

#### Practice

Use `infer` functions to create a 98% bootstrap confidence interval for the 
population *median* price of 1-bedroom apartments in Manhattan. Do this in a
single code chunk. Provide an interpretation of your result.

```{r practice_1}

```

### A confidence interval for the population proportion - $p$

Consider the mice radiation dataset. Suppose we want to compute a 95% confidence
interval for the proportion of mice that died while not on the treatment,
regardless of the dose.

First, identify the following:

- Population: 
- Parameter of interest:
- Sample:
- Sample size:

```{r boot_mice}
boot_prop <- mice %>% 
  filter(treat == 0) %>% 
  mutate(died = if_else(died == 1, "yes", "no")) %>% 
  specify(response = died, success = "yes") %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "prop")
```

```{r ci_mice}
ci_mice_died <- get_ci(boot_prop, level = .95)
ci_mice_died
```

What does this mean?

**Interpretation**:

```{r visualize_ci_mice}
visualise(boot_prop) +
  shade_ci(ci_mice_died)
```

#### Practice

Consider the mice radiation results. Suppose we want to compute a 90% confidence
interval for the proportion of mice that lived while on the treatment,
regardless of the dose.

```{r practice_2_bootstrap}

```



```{r practice_2_ci}

```



```{r practice_2_visualize}

```


Give an interpretation of your result.


# References

1. Source: C.W. Hammond, et al. (1955). "The Effect of Streptomycin Therapy
   on Mice Irradiated with Fast Neutrons", Radiation Research, Vol2,#4,
   pp.354-360

